---
title: "Data Manipulation With Dplyr"
author: "Subash"
output:
  pdf_document: default
  html_document: default
---

# Installing dplyr
```{r}
#install.packages("dplyr")
library(dplyr)
```


# Installing nycflights13
This data is an RDMS: a Relational Database Management System, it is made up of more than one table of data which are related to each other.
- flights
- airlines
- airport
- planes
- weather

```{r}
#install.packages("nycflights13)
library(nycflights13)
```


## Flights (Main Data)
Details of all flights in the year 2013
```{r}
head(flights)
```


## Airlines Names
```{r}
head(airlines)
```

## Airport Metadata
```{r}
head(airports)
```

## Planes Metadata
```{r}
head(planes)
```

## Weather (hourly)
```{r}
head(weather)
?weather
```

#  Grouping and Summarizing

```{r}
month_delay <- flights |> 
  group_by(month) |> 
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE))
```

```{r}
carrier_delay <- flights |> 
  group_by(carrier) |> 
  summarize(avg_carr_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_carr_arr_delay = mean(arr_delay, na.rm = TRUE))
```

# Arranging
```{r}
carrier_delay |> 
  arrange(desc(avg_carr_dep_delay))
```

# Filtering
```{r}
flights |> 
  filter(month == 1 & dep_delay < 0 & arr_delay < 0)

carrier_delay
```

# Selecting
```{r}
flights |> 
  select(!(hour:time_hour))
```

# Creating Variables
```{r}
carrier_speed <- flights |> 
  mutate(speed = distance/(air_time/60)) |> 
  select(carrier, speed) |> 
  group_by(carrier) |> 
  summarize(avg_speed = mean(speed, na.rm = TRUE)) |> 
  arrange(desc(avg_speed))
```


```{r}
flights |> 
  mutate(speed = distance/(air_time/60), .keep = "all")
```


# Renaming
```{r}
flights |> 
  rename(destination = dest)
```

# Mutating Joins
```{r}
top5_carrier_speed <- carrier_speed |> 
  head(5)
top5_carrier_speed
```

```{r}
full_join(x = top5_carrier_speed, 
           y = airlines, 
           by = "carrier")

inner_join(x = carrier_delay, 
           y = airlines, 
           by = "carrier") |> 
  select(carrier, name, avg_carr_dep_delay, avg_carr_arr_delay) |> 
  arrange(avg_carr_dep_delay) |> 
  head(5)
```

```{r}
inner_join(x = top5_carrier_speed, y = airlines, by = "carrier")
```

# Filtering Join
```{r}
anti_join(x = airlines,
          y = top5_carrier_speed,
          by = "carrier")
```